---
title: "p8105_hw2_hn2453"
author: "Huiyan Ni"
date: "2024-10-02"
output: github_document
---

## problem 1

```{r}
library(tidyverse)
library(readxl)
```

Import the data related to each entrance and exit for each subway station in NYC.
Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance.
```{r}
nyc_sub = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                   col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c"))

nyc_sub = nyc_sub |> janitor::clean_names() |> 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) |> 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

```
After importing original dataset, making route variables to character variable, selecting line, station_name, station_latitude, station_longitude, all route variables, entry, exit_only, vending, entrance_type, and ada columns, and convert the entry variable from character to a logical variable, now the dataset is 1868 * 20.   

How many distinct stations are there?
```{r}
nyc_sub |> 
  select(station_name, line) |> 
  distinct()
```
There are totally 465 rows, thus, there are 465 distinct stations.

How many stations are ADA compliant?
```{r}
nyc_sub |> 
  filter(ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```
There are totally 84 rows, thus, 84 distinct stations are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?
```{r}
nyc_sub |> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```
The proportion of station entrances without vending allow entrance is about 0.377.

Reformat data so that route number and route name are distinct variables.How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?
```{r}
nyc_sub |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()

nyc_sub |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A", ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()

```
60 distinct stations serve the A train, and 17 stations that serve the A train are ADA compliant.

## problem 2

Read and clean the Mr. Trash Wheel sheet. Specify the import sheet is "Mr. Trash Wheel", skip the first row due to the rows is non-data, define blank as na, convert variable names to reasonable variable names, omit rows that do not include dumpster-specific data(value is na), round the number of sports balls to the nearest integer and converts the result to an integer variable, and remove 2 useless columns producing after import data.
```{r}
mr_trash = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Mr. Trash Wheel",
                      col_names = TRUE,
                      skip = 1,
                      na = c("","NA")) |> 
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(sports_balls = as.integer(round(sports_balls)))
  
 mr_trash = mr_trash |> select(-x15,-x16)
 head(mr_trash)
```
Use a similar process to import, clean, and organize the data for Professor Trash Wheel and Gwynnda.
```{r}
pro_trash = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Professor Trash Wheel",
                      col_names = TRUE,
                      skip = 1,
                      na = c("","NA")) |> 
  janitor::clean_names() |>
  filter(!is.na(dumpster))
  
 head(pro_trash)
```
```{r}
gwy_trash = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Gwynnda Trash Wheel",
                      col_names = TRUE,
                      skip = 1,
                      na = c("","NA")) |> 
  janitor::clean_names() |>
  filter(!is.na(dumpster))

head(gwy_trash)
```

Add an additional variable to note different trash wheel before combining these data sets.
```{r}
mr_trash = mr_trash |> mutate(wheel_name = "Mr. Trash Wheel")
pro_trash = pro_trash |> mutate(wheel_name = "Professor Trash Wheel")
gwy_trash = gwy_trash |> mutate(wheel_name = "Gwynnda Trash Wheel")

```

Combining these data sets.
```{r}
mr_trash$year = as.numeric(mr_trash$year)
trash_wheel = bind_rows(mr_trash,pro_trash,gwy_trash)
trash_wheel
```
From the tibble, we have 1033 rows, and 15 variables: dumpster, month, year, date, weight_tons, volume_cubic_yard, plastic_bottles, polystyrene, cigarette_butts, glass_bottles, plastic_bags, wrappers, sports_balls, homes_powered, and wheel_name. 

What was the total weight of trash collected by Professor Trash Wheel?
```{r}
weight_pro = trash_wheel |> 
  filter(wheel_name == "Professor Trash Wheel") |> 
  pull(weight_tons) |> 
  sum(na.rm=TRUE)
```

The total weight of trash collected by Professor Trash Wheel is `r weight_pro` tons. 

What was the total number of cigarette butts collected by Gwynnda in June of 2022?
```{r}
cigarette_gwy = trash_wheel |>
  filter(wheel_name == "Gwynnda Trash Wheel") |>
  filter(month == "June") |>
  filter(year == "2022") |>
  pull(cigarette_butts) |>
  sum(na.rm=TRUE)

cigarette_gwy
```
The total number of cigarette butts collected by Gwynnda in June of 2022 is `r cigarette_gwy`.

## problem 3

Import and create three data sets from three different excels. 
```{r}
bake = 
  read_csv("data/gbb_datasets/bakes.csv", 
                      col_names = TRUE,
                      show_col_types = FALSE,
                      na = c("","NA","N/A")) |> 
  janitor::clean_names() 

head(bake)

baker = 
  read_csv("data/gbb_datasets/bakers.csv", 
                      col_names = TRUE,
                      show_col_types = FALSE,
                      na = c("","NA","N/A")) |> 
  janitor::clean_names() 

head(baker)

result = 
  read_csv("data/gbb_datasets/results.csv", 
                      col_names = TRUE,
                      show_col_types = FALSE,
                      skip = 2,
                      na = c("","NA","N/A")) |> 
  janitor::clean_names() 

head(result)

```
Finding the baker name in bakes.csv and results.csv is just first name, but the baker name in bakers.csv is full name. We need create a new variable for baker dataset in order to combine three data sets together. 
```{r}
baker$baker <- sapply(strsplit(baker$baker_name, " "), `[`, 1)
head(baker)
```

Create a single, well-organized dataset with all the information contained in these data sets.
```{r}
bake_result = full_join(bake, result, by = c("series","episode","baker"))

bake_total = full_join(bake_result, baker, by = c("baker","series"))
bake_total

```
Check the completeness of the final dataset.
```{r}
bake_na = anti_join(bake, bake_total, by = c("series","episode","baker"))
bake_na

baker_na = anti_join(baker, bake_total, by = c("series","baker"))
baker_na

result_na = anti_join(result, bake_total, by = c("series","episode","baker"))
result_na
```
All three datasets are null, thus, the completeness no problem. 


