---
title: "p8105_hw2_hn2453"
author: "Huiyan Ni"
date: "2024-10-02"
output: github_document
---

## problem 1

```{r}
library(tidyverse)
library(readxl)
```

Import the data related to each entrance and exit for each subway station in NYC.
Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance.
```{r}
nyc_sub = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                   col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c"))

nyc_sub = nyc_sub |> janitor::clean_names() |> 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) |> 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

```
After importing original dataset, making route variables to character variable, selecting line, station_name, station_latitude, station_longitude, all route variables, entry, exit_only, vending, entrance_type, and ada columns, and convert the entry variable from character to a logical variable, now the dataset is 1868 * 20.   

How many distinct stations are there?
```{r}
nyc_sub |> 
  select(station_name, line) |> 
  distinct()
```
There are totally 465 rows, thus, there are 465 distinct stations.

How many stations are ADA compliant?
```{r}
nyc_sub |> 
  filter(ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```
There are totally 84 rows, thus, 84 distinct stations are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?
```{r}
nyc_sub |> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```
The proportion of station entrances without vending allow entrance is about 0.377.

Reformat data so that route number and route name are distinct variables.How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?
```{r}
nyc_sub |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()

nyc_sub |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A", ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()

```
60 distinct stations serve the A train, and 17 stations that serve the A train are ADA compliant.

## problem 2

Read and clean the Mr. Trash Wheel sheet. Specify the import sheet is "Mr. Trash Wheel", skip the first row due to the rows is non-data, define blank as na, convert variable names to reasonable variable names, omit rows that do not include dumpster-specific data(value is na), round the number of sports balls to the nearest integer and converts the result to an integer variable, and remove 2 useless columns producing after import data.
```{r}
mr_trash = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Mr. Trash Wheel",
                      col_names = TRUE,
                      skip = 1,
                      na = c("","NA")) |> 
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(sports_balls = as.integer(round(sports_balls)))
  
 mr_trash = mr_trash |> select(-x15,-x16)
 head(mr_trash)
```
Use a similar process to import, clean, and organize the data for Professor Trash Wheel and Gwynnda.
```{r}
pro_trash = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Professor Trash Wheel",
                      col_names = TRUE,
                      skip = 1,
                      na = c("","NA")) |> 
  janitor::clean_names() |>
  filter(!is.na(dumpster))
  
 head(pro_trash)
```
```{r}
gwy_trash = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Gwynnda Trash Wheel",
                      col_names = TRUE,
                      skip = 1,
                      na = c("","NA")) |> 
  janitor::clean_names() |>
  filter(!is.na(dumpster))

head(gwy_trash)
```

Add an additional variable to note different trash wheel before combining these data sets.
```{r}
mr_trash = mr_trash |> mutate(wheel_name = "Mr. Trash Wheel")
pro_trash = pro_trash |> mutate(wheel_name = "Professor Trash Wheel")
gwy_trash = gwy_trash |> mutate(wheel_name = "Gwynnda Trash Wheel")

```

Combining these data sets.
```{r}
mr_trash$year = as.numeric(mr_trash$year)
trash_wheel = bind_rows(mr_trash,pro_trash,gwy_trash)
trash_wheel
```

```{r}

```

